Using Scipion for Stream Image Processing at Cryo-EM Facilities

### 2018-04-23 3.workflow_scilifelab.tex
I think this is done except for the cover letter

The facilities can be accessed by Swedish researchers through 
a peer-reviewed process. Applications can be submitted through an application portal (\url{https://cryoem.scilifelab.se/}). % and will be evaluated once every three months based on their scientific merit and technical feasibility by a national Project Evaluation Committee. %On the other side, some time of the facility is reserved for internal research groups at Stockholm University. 
Internally, a online booking system called \emph{Booked Scheduler} centralizes the reservations of the microscopes and other instruments.

At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion streaming processing. Before, there were some home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf) but the microscope operators recognized the importance of a more general framework. In particular, they were interested in the possibility of modifying easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection. Initially, the same setup script used at CNB was adopted, but it did not fully satisfy the facility requirements and a new one has been developed. In the following sections we will briefly describe the computational infrastructure at the \scilifelab node and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion. 
%[** thalos falcon III, titan k2 and falcon III]

\subsubsection{Network Setup and IT Infrastructure}
At the heart of cryo-EM National Facility computational infrastructure there is a storage and pre-processing server (the staging server). This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs (10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card. The storage pool is exported via NFS and Samba, so the microscope computers can write data there directly and other machines can access it.
%Unfortunately, the K2 [** Jose Miguel, what is K2?, do you have 2 cameras? in the same microscope or in different ones, this is confusing, comment when ] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server. %, in order to avoid interruption of data collection due to network issues.

In order to avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III. Hence an script to continuously move the produced data from the adequate storage to the staging server has been created.

@@

%ata on the staging server, in an organized folder structure, we
%can run pre-processing directly on the server.] If data processing is started soon
%after data collection the incoming data file should be in RAM-cache and  no
%disk access in needed, which is important with performance in mind. 
After processing, users may  copy their data to external USB drives from two workstations in the microscope room that have access to the storage pool via NFS.
There is also a dedicated download server which has access to the NFS export. This download server is accessible from  the Internet and allow downloads via rsync (over SSH) or SFTP.

@@

When a user starts a data collection, the wizard is executed using the command \textit{session-titan} (or \textit{session-talos}). The wizard determines which user has booked the microscope by fetching today's booking information from \emph{Booked Scheduler} through their web API. From the reservation information the wizard will also detect if it is an internal or a national facility project. 
In case of the latter, additional information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address). After gathering all this information the wizard will store it, create the processing environment and launch \scipion. %For example, the movies-import protocol will already point to the location where the files will be written for this session. Moreover, taking into account which microscope-camera is being  used, the files pattern can be inferred.

@@

The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}. It has been designed to satisfy the requirements for \scilifelab and, although readers are welcomed to try it, no support is offered.  \emph{Booked Scheduler} is available at (\url{https://www.bookedscheduler.com/})

>>>>

The facilities can be accessed by Swedish researchers through 
a peer-reviewed process. Applications can be submitted through an application portal (\url{https://cryoem.scilifelab.se/}). % and will be evaluated once every three months based on their scientific merit and technical feasibility by a national Project Evaluation Committee. %On the other side, some time of the facility is reserved for internal research groups at Stockholm University. 
Internally, an online booking system called \emph{Booked Scheduler} centralizes the reservations of the microscopes and other instruments.

At the beginning of 2016, the Cryo-EM facility at \scilifelab became one of the early adopters of \scipion streaming processing. Although there already were using home-made scripts to perform motion correction (with motioncor2) and CTF estimation (with Gctf), the microscope operators recognized the importance of a more general
framework. In particular, they were interested in the possibility of modifying easily the processing workflow as well as provide users and staff with  graphical tools for data analysis and quick feedback regarding the data collection.
Although initially the same setup script used at
CNB was adopted, a new one was developed to fully satisfy the facility
requirements. In the following sections we will briefly describe the computational infrastructure at the \scilifelab node and the implementation of a Session Wizard to automate the initial setup while fetching information from other sources external to \scipion. 
%[** thalos falcon III, titan k2 and falcon III]

\subsubsection{Network Setup and IT Infrastructure}
A storage and pre-processing server (the staging
server) constitutes the core of the SciLifeLab computational infrastructure.
 This machine has roughly 200 TB of storage (4 ZFS RAIDZ2 pools with 11 HDDs each), 2 NVIDIA GeForce GTX 1070 GPUs, 2 Intel Xeon E5-2630v4 CPUs (10 cores, 2.2 GHz, HT), 384 GB RAM, and a dual-port 10 GbE network card. The storage pool is exported via NFS and Samba. The microscope computers can thus write data directly there and other machines can access it.

 The storage pool is exported via NFS and Samba, so the microscope computers can write data there directly and other machines can access it.
%Unfortunately, the K2 [** Jose Miguel, what is K2?, do you have 2 cameras? in the same microscope or in different ones, this is confusing, comment when ] computer writes data to a local SSD RAID while the Falcon-III does it to a dedicated storage server. %, in order to avoid interruption of data collection due to network issues.

To avoid interruption of data collection due to network issues each microscope has its own dedicated storage, that in the case of the Titan microscope is double since it may be connected to two different cameras, a K2 and a Falcon III. Hence an script to continuously move the produced data from the adequate storage to the staging server has been created.

@@

%ata on the staging server, in an organized folder structure, we
%can run pre-processing directly on the server.] If data processing is started soon
%after data collection the incoming data file should be in RAM-cache and  no
%disk access in needed, which is important with performance in mind. 
After processing, users may  copy their data to external USB drives from two workstations in the microscope room that have access to the storage pool via NFS.
There is also a dedicated download server which has access to the NFS export. This download server is accessible from  the Internet and allows downloads via rsync (over SSH) or SFTP.

@@

When a user starts a data collection, the wizard is executed using the command \textit{session-titan} (or \textit{session-talos}). The wizard determines which user has booked the microscope by fetching today's booking information from \emph{Booked Scheduler} through their web API. From the reservation information the wizard will also detect if it is an internal or a national facility project. 
In case of the latter, additional information will be retrieved from the Application Portal (such as principal investigator, project code, contact person or invoice address). After gathering all this information the wizard stores it, creates the processing environment and launches \scipion. %For example, the movies-import protocol will already point to the location where the files will be written for this session. Moreover, taking into account which microscope-camera is being  used, the files pattern can be inferred.

@@

The described session wizard can be found at  \url{https://github.com/delarosatrevin/scipion-session}. You are welcomed to download and customize it to meet your organization specific requirements. \emph{Booked Scheduler} is available at (\url{https://www.bookedscheduler.com/})

### 2018-04-23 4.future.tex
I think this is done except for the cover letter

Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, Relion, ...) at the time of the release. Therefore, an update of any of the integrated EM packages immediately after the \scipion release won't be available until next \scipion release (we are aiming for one major release per year). To decouple EM-package releases from \scipion releases, we are working on making the wrappers for EM packages  independent from \scipion. Plans for ensuring this independence include: (1) wrapper installation through pip (\url{https://pip.pypa.io/en/stable/}), (2) reimplementation of the wrappers as plug-ins that will be automatically detected and added to the application menu and (3) develop of standalone IO  and visualization libraries that make \scipion fully independent from other EM packages.

>>>>

Currently, \scipion includes the appropriate code (python wrappers) to talk to the integrated EM packages (Xmipp, Eman, Relion, ...) at the time of the release. Therefore, an update of any of the integrated EM packages immediately after the \scipion release will not be available until next \scipion release (we are aiming for one major release per year). To decouple EM-package releases from \scipion releases, we are working on making the wrappers for EM packages  independent from \scipion. Plans for ensuring this independence include: (1) wrapper installation through pip (\url{https://pip.pypa.io/en/stable/}), (2) reimplementation of the wrappers as plug-ins that will be automatically detected and added to the application menu and (3) develop of standalone IO  and visualization libraries that make \scipion fully independent from other EM packages.

### 2018-04-23 5.conclusions.tex
I think this is done except for the cover letter

 \scipion has been deployed in cloud and there is a public image, with all the required preinstalled cryoEM software,  available at Amazon WebServices EC2 and EGI federated clouds. These images make use of \scipion streaming capabilities, that is, they are able to process data as soon as it has been transfered to an Amazon storage unit, and can be used to handle  production peaks.

\scipion is provided 
freely as open source software. Online documentation
describing \scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively. Cloud installation instruction are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}. Other public software mentioned in this article as \emadmin, \emph{wizard session} and the different scripts used at eBIC are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, \url{\url{https://github.com/delarosatrevin/scipion-session}} and \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel} respectively.
If you would like to have more information about how to install \scipion or you have suggestions on how to improve it, contact us at scipion@cnb.csic.es.

>>>>

 \scipion has been deployed in cloud and there is a public image, with all the required preinstalled cryoEM software,  available at Amazon WebServices EC2 and EGI federated clouds. This image makes use of \scipion streaming capabilities, that is, they are able to process data as soon as it has been transfered to an Amazon storage unit, and can be used to handle  production peaks.

\scipion is provided 
freely as open source software. Online documentation
describing \scipion download and installation is available
at \url{http://scipion.cnb.csic.es/m/download_form/} and \url{https://github.com/I2PC/scipion/wiki/How-to-Install}, respectively. Cloud installation instructions are available at \url{https://github.com/I2PC/scipion/wiki/Scipion-in-the-Cloud}. Other public software mentioned in this article as \emadmin, \emph{wizard session} and the different scripts used at eBIC are available at \url{https://github.com/rmarabini/webservices/tree/master/EMadmin}, \url{https://github.com/delarosatrevin/scipion-session} and \url{https://github.com/I2PC/scipion/tree/release-1.1-headless-devel} respectively.
\scipion download statistics, usage map and protocol usage ranking are available at \url{http://scipion.i2pc.es/downloadstats}, \url{http://scipion.i2pc.es/report_protocols/scipionUsage/} and \url{http://scipion.i2pc.es/report_protocols/protocolTable/} respectively.

 For more information about how to install Scipion or
suggestions to improve it, contact us at scipion@cnb.csic.es.
